Weather-Related Damages to Life and Property in the United States: a Retrospective Analysis from National Weather Service Data Between 1950 and 2011

### 2021-03-28 Manuscript.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)

suppressMessages(library(tidyverse))
suppressMessages(library(dtplyr))
suppressMessages(library(data.table))
# library(skimr)
```

@@

## Introduction

We analyze this, that and something else...

@@

inv_crop_exp <- filter(temp_data, CROPDMG != 0, crop_mult == -999999) %>%
  select(CROPDMG, CROPDMGEXP, crop_mult, BGN_DATE) %>%
  as.data.frame() %>%
  nrow()

# Clean up
# rm(temp_data)
```

@@

```{r inj-fat-describe}

skimr::skim(mdata$INJURIES)
cat("\n\n\n")
skimr::skim(mdata$FATALITIES)

@@

# Make a copy of data and cast as DT
sel_data <- as.data.table(lean_data)
# Apply final transformations to DT
sel_data2 <- sel_data[, .(
    counter = .N, # Count elements in group
    # Total monetary damages per group
    total_dmg = sum(tot_dmg, na.rm = TRUE),
    # Total fatalities
    fatalities = sum(FATALITIES, na.rm = TRUE),
    # Injuries
    injuries = sum(INJURIES, na.rm = TRUE)
  ),
  # Grouping by event type
  by = .(event)][
    ,
    # Compute a "percent rank": rank events by number of
    # occurrences (desc), then assign a percentile by
    # dividing rank by denominator (i.e. nrows)
    # Note double dot syntax to access calling scope
    `:=`(perc_rank = frankv(counter)/.N,
         # Same for monetary damage
         dmg_rank = frankv(total_dmg)/.N,
         # And life damage
         inj_rank = frankv(injuries)/.N,
         fat_rank = frankv(fatalities)/.N
    )
  ][
    # Filter for frequency rank >=95th
    perc_rank >= 0.80
  ][
    order(-counter)
  ]

@@

```{r factor-loops}
suppressWarnings(
    # For every new event category
    for (i in seq(1:length(new_fcts))) {
        # Prepare storage array
        old_instances <- c()
        k <- 1
        # Loop over elements in list element
        for (k in seq(1:length(new_fcts[[i]]))) {
            # Match old event instances
            old_instances <- c(old_instances,
                               str_subset(sel_data2$event, new_fcts[[i]][k])
                                )
            k <- k + 1
        }
        # Add to list
        old_events[[i]] <- unique(old_instances)
        i <- i + 1
    }
)
# Name (tag) list elements
names(old_events) <- names(new_fcts)

# TODO: switch to using list names instead of manual entry of new factor levels
sel_data2$event2 <- fct_collapse(
  sel_data2$event,
  hail = old_events$hail,
  tornado = old_events$tornado,
  tsunami = old_events$tsunami,
  stormy = old_events$stormy,
  lightning = old_events$lightning,
  flood = old_events$flood,
  cold = old_events$cold,
  heat = old_events$heat,
  windy = old_events$windy,
  maritime = old_events$maritime,
  fog = old_events$fog,
  drought = old_events$drought,
  fire = old_events$fire,
  landslide = old_events$landslide,
  other_level = "Other events"
) %>%
  # Adjust spelling/capitalization of factor levels
  fct_recode(
    "Hail" = "hail",
    "Winds" = "windy",
    "Tornadoes" = "tornado",
    "Floods" = "flood",
    "Lightnings" = "lightning",
    "Cold weather" = "cold",
    "Storms/Rains" = "stormy",
    "Oceanic events" = "maritime",
    "Fires" = "fire",
    "Droughts" = "drought",
    "Hot weather" = "heat",
    "Fog" = "fog",
    "Landslides" = "landslide",
    "Tsunamis" = "tsunami"
  ) %>%
  # Reorder according to frequency, desc
  fct_infreq()

@@

```{r memory-release-cleanup}

>>>>

```{r setup}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)

suppressMessages(library(tidyverse))
suppressMessages(library(dtplyr))
suppressMessages(library(data.table))
```

@@

@@

inv_crop_exp <- filter(temp_data, CROPDMG != 0, crop_mult == -999999) %>%
  select(CROPDMG, CROPDMGEXP, crop_mult, BGN_DATE) %>%
  as.data.frame() %>%
  nrow()
```

@@

```{r inj-fat-describe}
skimr::skim(mdata$INJURIES)
skimr::skim(mdata$FATALITIES)
```

@@

# Make a copy of data and cast as DT
sel_data <- as.data.table(lean_data)
# Apply final transformations to DT
sel_data2 <- sel_data[, .(
    counter = .N, # Count elements in group
    # Total monetary damages per group
    total_dmg = sum(tot_dmg, na.rm = TRUE),
    # Total fatalities
    fatalities = sum(FATALITIES, na.rm = TRUE),
    # Injuries
    injuries = sum(INJURIES, na.rm = TRUE)
  ),
  # Grouping by event type
  by = .(event)][
    ,
    # Compute a "percent rank": rank events by number of
    # occurrences (desc), then assign a percentile by
    # dividing rank by denominator (i.e. nrows)
    # Note double dot syntax to access calling scope
    `:=`(perc_rank = frankv(counter)/.N,
         # Same for monetary damage
         dmg_rank = frankv(total_dmg)/.N,
         # And life damage
         inj_rank = frankv(injuries)/.N,
         fat_rank = frankv(fatalities)/.N
    )
  ][
    # Filter for frequency rank >=95th
    perc_rank >= 0.80
  ][
    order(-counter)
  ]
```

@@

```{r factor-loops, message=FALSE, warning=FALSE}
suppressWarnings(
    # For every new event category
    for (i in seq(1:length(new_fcts))) {
        # Prepare storage array
        old_instances <- c()
        k <- 1
        # Loop over elements in list element
        for (k in seq(1:length(new_fcts[[i]]))) {
            # Match old event instances
            old_instances <- c(old_instances,
                               str_subset(sel_data2$event, new_fcts[[i]][k])
                                )
            k <- k + 1
        }
        # Add to list
        old_events[[i]] <- unique(old_instances)
        i <- i + 1
    }
)
# Name (tag) list elements
names(old_events) <- names(new_fcts)

# TODO: switch to using list names instead of manual entry of new factor levels
sel_data2$event2 <- fct_collapse(
  sel_data2$event,
  hail = old_events$hail,
  tornado = old_events$tornado,
  tsunami = old_events$tsunami,
  stormy = old_events$stormy,
  lightning = old_events$lightning,
  flood = old_events$flood,
  cold = old_events$cold,
  heat = old_events$heat,
  windy = old_events$windy,
  maritime = old_events$maritime,
  fog = old_events$fog,
  drought = old_events$drought,
  fire = old_events$fire,
  landslide = old_events$landslide,
  other_level = "Other events"
) %>%
  # Adjust spelling/capitalization of factor levels
  fct_recode(
    "Hail" = "hail",
    "Winds" = "windy",
    "Tornadoes" = "tornado",
    "Floods" = "flood",
    "Lightnings" = "lightning",
    "Cold weather" = "cold",
    "Storms/Rains" = "stormy",
    "Oceanic events" = "maritime",
    "Fires" = "fire",
    "Droughts" = "drought",
    "Hot weather" = "heat",
    "Fog" = "fog",
    "Landslides" = "landslide",
    "Tsunamis" = "tsunami"
  ) %>%
  # Reorder according to frequency, desc
  fct_infreq()
```

@@

```{r memory-release-cleanup, include=FALSE}
